---
title: "Coolest Thing in R"
date: '2022-06-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.path = "../images/")
```

## R vs. Python
I was first exposed to R, followed by SAS and Python. And I am more proficient in R, as compared to the others. I like the consistency of R and its power of visualizing the data. 
```{r }
plot(iris)
```

For example, the ggplot2 library provides  a great deal of intuitive customizability. 
While R emphasizes statistical analysis, Python gives me the necessary skills to work on business,digital products, open-source projects and various application development outside of data science. Due to its terminal function, Python is also a great tool for bioinformatic analysis, such as gene sequencing quality checks and trimming, alignment to reference genome and DEG calling. Those functions are not included in R. Since R is an open source language, it is the users' responsibility to validate and manage the version of the packages. I believe it's not hard to learn R, because it was my first language to learn and I got used to its logics. And as an advanced language, coding is much easier than other languages, such as C++.  

