---
title: "ST558_Blog3"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(easypackages)
libraries("httr", "tidyverse", "jsonlite", "DT", "stringr", "RColorBrewer")
```

One of the coolest thing I've learned about programming in R is its powerful packages, which exist to do almost anything. For example, I'd like to show the coronavirus cases by country in the world map. I can just read in the world map data set, which includes the longitude, latitude and country names. 

```{r}
# Read in the world map
mapdata <- map_data("world")
head(mapdata)
```

Then, I read in the coronavirus cases from API. 

```{r, include= FALSE}
# Write a function to read in the data set from API
get_Data <- function(url){
                myData<- GET(paste0("https://api.covid19api.com", url))
                parsed <- fromJSON(rawToChar(myData$content))
                return(parsed)
}
```

```{r}
covid_summary <- get_Data("/summary")$Countries %>% as_tibble()
```

```{r, include=FALSE}
countries_Names <- function(countries){
  ifelse(countries %in% c("Antigua and Barbuda", "Brunei Darussalam", "Iran, Islamic Republic of", "Micronesia, Federated States of", "Taiwan, Republic of China", "Tanzania, United Republic of", "Trinidad and Tobago"), new_countries <- str_replace(word(countries, 1), ",", ""), 
         ifelse(countries == "Congo (Brazzaville)",  new_countries <- "Republic of Congo", 
                ifelse(countries == "Congo (Kinshasa)",  new_countries <- "Democratic Republic of the Congo",
                       ifelse(countries == "Korea (North)", new_countries <- "North Korea",
                              ifelse(countries == "Lao PDR", new_countries <- "Laos",
                                     ifelse(countries == "Macedonia, Republic of", new_countries <- "North Macedonia", 
                                            ifelse(countries == "Republic of Kosovo", new_countries <- "Kosovo", 
                                                  ifelse(countries == "Saint Vincent and Grenadines", new_countries <- "Saint Vincent",
                                                         ifelse(countries == "Syrian Arab Republic (Syria)", new_countries <- "Syria",
                                                                ifelse(countries == "United Kingdom", new_countries <- "UK",
                                                                       ifelse(countries == "United States of America", new_countries <- "USA",
                                                                              ifelse(countries == "Viet Nam", new_countries <-"Vietnam", 
                                                                                     ifelse(countries == "Saint Kitts and Nevis", new_countries <- "Saint Kitts", 
                                                                                            ifelse(countries == "Korea (South)", new_countries <- "South Korea",
                                                                                                   ifelse(countries == "Russian Federation", new_countries <- "Russia", 
                                                                                                          new_countries <- countries)))))))))))))))
}


# Using the countries_Names function to create a new variable called new_Countries to match the country names in the map data set
covid_summary$new_Countries <- countries_Names(covid_summary$Country)
```

After combine the coronavirus and world map data sets. I can simply create a worldwide coronavirus map by ggplot.  

```{r}
# Combine the map and covid_summary data sets 
mapData <- left_join(mapdata, covid_summary, by = c("region" = "new_Countries"))

# Remove the rows which do not contain total confirmed cases numbers 
mapdata1 <- mapData %>% filter(!is.na(mapData$TotalConfirmed))
```


```{r}
ggplot(mapdata1, aes(x= long, y =lat, group = group)) +
  geom_polygon(aes(fill = TotalConfirmed), color = "darkgreen") + 
  scale_fill_gradient(name = "Cases", low='#EEEEEE', high='darkgreen', na.value = "grey") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        rect = element_blank()) +
  labs(title = "Coronavirus Worldwide")
```

This map shows us that the United States of America has the largest total confirmed cases, followed by India and Brazil.